The Chemistry Development Kit (CDK). 3. Atom typing, Rendering, Molecular Formula, and Substructure Searching

### 2016-08-24 article.tex
Update CTfile (molfile) section and include relevant citations.

  \subsubsection*{MDL molfile improvements}

  TODO: S group, V3000 formats, ...

>>>>

\subsubsection*{CTfile Format Improvements}

The molfile format is still very popular, despite it being a proprietary
format it has become a de facto standard. The format forms the core of the larger
CTfile family which was originally developed by MDL Information Systems \cite{Dalby92}. The
current format specification is published by BIOVIA and available on 
request \cite{ctfilespec}.

%% CTfile Format Support
%% To tabulate, r=read, w=write
%% MOLfile V2000=rw, V3000=rw
%% RXNfile V2000=rw, V3000=r
%% SDfile MOLfile V2000=rw, MOLfile V3000=r, RGfile=n/a
%% RGfile V2000=rw
%% RDfile n/a

The CTAB block (connection table) of a molfile comes in two versions, V2000
and V3000. The V3000 provides several enhancements including but not
limited to: removing atom and bond count limits, enhanced stereochemistry,
and link nodes. For backwards compatibility V2000 is often preferred resulting
in limited usage of V3000.

CDK \textbf{2.0} adds support for V3000 and has optimized and extended
support for V2000. Currently these are considered separate formats requiring
a user to know what version is being read beforehand. Future APIs will aim
to simplify this and provide a unified reader.

%% TODO(JWM): Sgroup Examples

CTfile Sgroups capture and organise high level information about sets of atoms
and bonds \cite{Gushurst91}. There are four types of Sgroup: Display Shortcuts, Polymers,
Mixtures, and Data. The most familiar Sgroups from an end user perspective are structure 
repeat units (e.g. bracketing) and abbreviations (figure: ??). CDK \textbf{2.0} adds 
supports for representation, reading, writing, and depiction.

% note, I pie chart would be cool, showing the various relative amounts of the
% various formats.
